On Devon Allen's Disqualification at the 2022 Track and Field World Championship

### 2023-04-25 Manuscript/manuscript.tex
comments through Section 2

%1)
Devon Allen, an alumnus of the University of Oregon in 
Eugene, Oregon, was
expected to be a contender for the 2022 World Track and Field Championships 110 
meter hurdle.  After running the event in 12.84 seconds in early June 2022 
(only 0.04 seconds away from the world record \citep{wa2022preview}) and coming 
in third in the event at the United States Track and Field 
Championships also in Eugene at the end of June, Allen was poised to 
compete in the 2022 World Track and Field Championships again in Eugene in August.
%After coming in third in the 110 meter hurdles at the United States Track and 
%Field Championships also in Eugene at the end of June, Allen was poised to 
%compete in the 2022 World Track and Field Championships also 
%located at Eugene in August.
%two months later. 
At the World Championships, Allen made it through the 
heats and semifinals to reach the finals while posting reaction times of 0.123
in the heats \eds{if more than one heat, include all of them or the average;
if only one heat, change to singular form}
\of{There are roughly 5 heats, but each athlete only competes in one heat} 
and 0.101 in the semifinal.  
%Allen, who attended the University of Oregon also in Eugene, 
%(the location of US World Championship 
%\eds{US World? remove World?} and the World Championship) 
%was
%considered to be a favorite for the 110 meter hurdle after running 12.84 seconds
%in early June; only 0.04 seconds away from the world record \citep{wa2022preview}.
%Thus 
His disqualification by reacting in 0.099 seconds instead of 0.1 was met
with shock from both Allen and fans alike.  NBC, who was broadcasting the
championships in the United States, later uploaded a video to their YouTube
channel that showed the disqualification; it can be found here:
\url{https://www.youtube.com/watch?v=D6NXTMo-1yM}. The crowd erupted with jeers 
when they found out that Allen was disqualified, angry at the result.  Had Allen 
reacted just 0.001 seconds slower he would have been allowed to compete and been 
spared waiting another two years until the next World Track and Field Championship.

\eds{final two sentence seems out of place here; consider moving elsewhere?}
\of{Moved to the end of paragraph 4 of introduction}

%2)
Allen's disqualification has been discussed extensively in online communities
such as \url{www.LetsRun.com}, a website that is part message board and part
news. The message board,
which functions similarly to Reddit is very active during important running
events, including during the finals of major events such as
the Olympics and the World Track and Field Championships.  However, it is the
news section of the website that generated articles such as \citet{johnson2022was}
and \citet{johnson2022data} from  Robert Johnson, who created and runs 
the LetsRun website.  He argued that there was an error
with the timing equipment at the 2022 Track and Field World Championship and
cited graphs, simple statistics such as median reaction time, and compared the 
reaction times at the 2022 World Track and Field Championship to other competitions
such as the USA Track and Field Championships and other World Championships.
However, their analysis lacked formal analysis. The alleged 99.95\% chance Allen
was wrong came from that the lowest average reaction time for the men's 100 meter,
men's 110 meter hurdle, and women's 100 meters from 2003 to 2022 all occurred at
the 2022 World Championships and Johnson claimed that the probability of this
happening was $\frac{1}{13} \cdot \frac{1}{13} \cdot \frac{1}{13} = 0.05\%$
\citep{johnson2022was}.
\eds{compared the reaction times at the 2022 World Track and Field  Championship 
to what? Also restate here that no formal statistical analyses have been done.}
\of{Cleaned this up a little but still a little messy.}

@@

\jy{Could be condensed}
\citet{pilianidis2012start} found
significant differences between reaction times at several world championships
between 1997 and 2009 for the men's 110 meter hurdles.  Another important
conclusion was that reaction times did not decrease over the twelve years
covered by the research in the study \citep{pilianidis2012start}. This contradicts
what was found in a 2021 study that concluded the year of competition has a
significant effect on reaction times \citep{zhang2021correlation}.  Their
analysis looked at data from the 2011-2019 World Championships across multiple
sprint events and concluded that from 2011 to 2019, reaction times got
faster. However, not all studies have been in favor of a decrease in the
reaction time barrier. \citet*{brosnan2017effects} argue for increasing 
the reaction time barrier to 0.115 seconds, and cited evidence from European 
Championships and World Championships in their article.  A study looking 
at the 2008 Olympics likewise supported the conclusion that athletes could not
react in as little as 0.1 seconds \citep{lipps2011implications}.  Another study
concluded that visual reaction times to the starting gun may be faster by 0.007
seconds than the IAAF method, which is from force on block starts 
\citep{holmes2018method}.  While this may seem trivial, Allen was disqualified
only by 0.001 seconds. It is also worth noting that one study looking at data 
from various IAAF meets found that the fastest reaction times for men were found
between the ages of 26 and 29 \citep{tonnessen2013reaction}.  This is notable 
because Allen was 27 years old at the 2022 Track and Field World Championship 
and in his physical peak.

@@

%4)
\jy{language needs to be formalized and condensed.}
The objective to this paper is to analyze the reaction times at the 2022
World Track and Field Championship to check for any abnormalities, and we approach
this two different ways.  First we try to model the data, which is
the reaction times of every non-disqualification since the 1999 World Championships. 
The model we use is a gamma linear mixed-effects model with a random effect
accounting for the venue (or year) effect. 
Then we compare reaction times for athletes who have competed
at additional competitions to the 2022 World Track and Field Championship.  By
comparing the same athlete across multiple competitions we are able to see if the
2022 data differed significantly from other competitions.
Our investigation into reaction times is important because reaction times have
been shown to affect the overall time of the race \citep{delalija2008reaction}.
Thus, faster reaction times lead to faster overall times; which is the goal of
every sprinter who steps onto the track.
%\eds{I recommend moving the rest of this paragraph to Section 2.2 as I don't
%think it is critical for readers to know all the details here.}
%First we can compare the reaction times of athletes who
%competed at the World Track and Field Championship in both 2019 and 2022, and 
%second we can compare how athletes performed at the 2022 championship relative 
%to their national/qualifiers meet. In order to reduce any chances
%of athletes improving reaction time, we can also look at data from qualifying meets.
%Almost every athlete had to qualify by proving they are elite at the national level, 
%usually by coming in at least fourth. Thus, data was compiled on athletes who competed
%at the 2022 World Track and Field Championships on what these athletes also ran 
%at their country's national championship. If there was any argument that athletes
%have improved their reaction time from 2019 to 2022, it should now be considered
%nearly null, as the national competitions took place between June and July 2022 
%and the World Championships were in August. The equipment used to measure reaction
%time cannot be guaranteed to be the same from national to international
%competition, but the time span between meets has been reduced considerably.

The rest of this paper is organized as follows. Section~\ref{sec:Data} describes 
how data was collected and begins to detail the generalized linear mixed-effect 
(GLMM) model that is further developed in Section~\ref{sec:Methods}. The 
signed-rank test that analyzes reaction times for the same athletes is
also described in Section~\ref{sec:Methods}.  Section~\ref{sec:Results}
 provides the results of the two analyses and ends 
by questioning the reaction time barrier.  Lastly, Section~\ref{sec:Discussion}
discusses the impact and limitations of the paper.
\eds{capitalize first letter when referring to a numbered section, table, figure, 
etc. ``Section 4 provides...'' vs. ``The next section provides...''}

@@

There are two types of data that are explored throughout this paper: data from
1999-2022 for every world championship, and data from athletes who competed
the 2022 World Championships and another competition (either 2019 World 
Championship or a 2022 national competition held within two months of the  
2022 World Championships). The first data set is used for the GLMM portion
of the paper, while the second data set is used for the rank-based comparisons.

@@

Data was taken from the World Athletics and covers the men's 110 meter hurdles 
and the women's 100 meter hurdles from 1999 to 2022.  The variable of prediction 
interest is reaction time, and the predicting variables are year, stage of 
competition (heats, semifinal, final), total time, and gender.  
The data can be found in the appendix~\ref{sec:Appendix}. Factors such as name, 
country, and identification number were not considered important for the 
purposes of this research besides the times for Devon Allen and the other 
United States athletes mentioned above.  The variable ``Stage"
refers to whether the observation occurred during the ``heats" (Preliminary round) which
is denoted by ``H" in the data, ``S" for ``semifinals" (Of which there are either 2 or 3 
heats each year), and ``F" for finals. The variable ``Gender" has levels ``M" and ``F" 
which stand for ``male" and ``female" respectively.  The variable ``Batch" takes a 
numerical value and every heat of every round from 1999 to 2022 was assigned it's
own number.  This was added because of suspected issues that reaction time may be
correlated with other runners reacting quickly (one runner reacting quickly may
result in other runners reacting quickly).  There is no statistical significance to
how the batches were labeled, they were ordered chronologically by meet but
reverse chronologically by year (Batch 1 is a preliminary Heat of 2022 and Batch
111 corresponds to the 1999 finals). Included are
box plots of the reaction times of the men's finals and the men's semifinals and
finals pooled together.  We will refer to the latter as the ``pooled data"
throughout the paper.  The heats or preliminary rounds were not studied because
athletes react faster as the competition progresses.  Athletes react faster in
the finals than the semifinals which are faster than the heats 
\citep{zhang2021correlation}.  If we want to determine if Allen's reaction time
is reasonable, then we thus need to look at the men's final data, with the men's
pooled data being included in order to increase the sample size.  In 2022, there
were only five data points due to two disqualifications and one athlete who 
qualified but did not compete.
The data that looked at in the paper is best visualized using a box plot as show 
in Figure~\ref{fig:Boxplots}.  It is clear that the 2022 box plot for both the
men's finals and semifinals data is lower (indicating a faster reaction time)
than in previous years.  The median for the men's pooled data in 2022 was 0.129
and the median for the men's final data in 2022 was 0.126.  For comparison,
Brosnan found reaction times of 0.156 seconds for men when looking at data from
1999 to 2014 \citep{brosnan2017effects}.

Something that is extremely important to note in the data is the inconsistency 
of World Athletics in terms of how they penalize false starts.  From 2007--2009,
the IAAF (the former name for World Athletics, and the governing body for the 
World Track and Field Champions) instituted a rule change that allowed one 
warning false start before a sprinter was disqualified \citep{iaaf2009falsestart}  This 
essentially gave sprinters a second chance and reduced the penalty for a false 
start.  There were 18 male and 7 female false starts at the 2007 World 
Championships, 18 male and 7 female starts at the 2009 World Championships. 
Starting in 2011, the rule was scrapped, and the old policy which automatically 
disqualifies runners who false start was reinstituted.  The effect? At the 2011 
World Championship there were 6 male disqualifications and 4 female 
disqualifications \citep{iaaf2009falsestart}. By returning to the harsher policy and 
cracking down on false starts, World Athletics had reduced the number of 
false starts by two thirds. One study that looked at different false start
rules that IAAF has imposed from 1997 to 2009 and found that the reaction 
times when there was less leniency improved by a statistically significant amount
\citep{haugen2013effect}. This was desirable for World Athletics as false starts can 
make already long track meets more tedious, both for athletes and viewers watching the 
television broadcast.  Since 2011, IAAF has instituted the one false start rule
\citep{iaaf2009falsestart}.  However, we chose not to include women's data in this
paper due to the numerous papers citing difference in reaction times in men and
women's reaction times: \citep{lipps2011implications, babicc2009reaction,
panoutsakopoulos2020gender}

\subsection{Beyond World Championship}
This section was motivated from a qualitative analysis that started by looking
at how United States athletes reacted at the 2022 United States Track and Field
Championships. From June 23-26 2022, USA held its Track
and Field Championship to decide who to send to the World Championships at Hayward 
Field, the exact same venue as used for the World Track and Field Championship a month 
later.  Thus, a baseline is able to be established for the four USA athletes who 
competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
Grant Holloway, and Devon Allen. Every athlete reacted faster in all the World 
Track and Field Championship races compared to the USA Track and Field 
Championship races. Devon Allen raced three times at the USA Track and
Field Championship and three times at the World Track and Field  Championship. 
His reaction times were significantly faster in the World Track and
Field Championship compared to the USA Track and Field Championship as were his
teammates.  This comparison suggests that the timing system used between the two
meets may have been meaningfully different to the point where Devon Allen was
disqualified more because of faulty equipment rather than because he reacted
too quickly.  Table~\ref{fig:USAvsWorld} highlights the difference in reaction times for
Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. ``USA" denotes
the USA Track and Field Championships, ``World" denotes the World Track and Field
Championships,   All numbers listed in the table are the reaction
time for the respective athletes in seconds.

\begin{table}
\begin{center}
  \caption{Rection times for four United States Athletes at the 2022 USA Track
  and Field World Championships and the 2022 World Track and Field Championship.
  ``N/A" was used for any athlete that did not compete and thus did not register 
  a reaction time.}
  \begin{tabular}{c c c c c c c} 
   \toprule
   Athlete & USA H & USA S & USA F & World H & World S & World F \\ [0.5ex] 
   \midrule
   Devon Allen & 0.201 & 0.153 & 0.160 & 0.123 & 0.101 & 0.099 \\ 
   Trey Cunningham & 0.186 & 0.185 & 0.182 & 0.115 & 0.120 & 0.109 \\
   Grant Holloway & 0.192 & 0.190 & N/A & 0.147 & 0.128 & 0.124 \\
   Daniel Roberts & 0.181 & 0.200 & 0.183 & 0.179 & N/A & N/A \\ [0.5ex]
   \bottomrule
  \end{tabular}
  \label{fig:USAvsWorld}

@@

\eds{unless target journal has specified otherwise, Table captions should go
above the table. (Figure captions go below the figure)}
\of{Moved table caption to be above}

This dataset is quite small to perform any formal statistical analysis on, but
the data was expanded to include data for athletes who competed at national
competitions (many of which were held from May-July of 2022).  Once the data
was collected, each athlete was designated to be a cluster and the stage of
competition (national or international) was specified to be the group.

@@

The World Championship data and the data beyond the World Champisonship were
analysed with GLMM and rank-based approaches, respectively.

>>>>

%1)
Devon Allen, an alumnus of the University of Oregon in 
Eugene, Oregon, was
expected to be a contender for the 2022 World Track and Field Championships 110 
meter hurdle.  After running the event in 12.84 seconds in early June 2022 
(only 0.04 seconds away from the world record \citep{wa2022preview}) and coming 
in third at the United States Track and Field 
Championships also in Eugene in late June, Allen was poised to compete in the 
2022 World Track and Field Championships at the same Eugene venue in August.
%After coming in third in the 110 meter hurdles at the United States Track and 
%Field Championships also in Eugene at the end of June, Allen was poised to 
%compete in the 2022 World Track and Field Championships also 
%located at Eugene in August.
%two months later. 
At the World Championships, Allen made it through the preliminary
heats and semifinals to reach the final heat while posting reaction times of 
0.123 in his preliminary heat and 0.101 in his semifinal.  
%Allen, who attended the University of Oregon also in Eugene, 
%(the location of US World Championship 
%\eds{US World? remove World?} and the World Championship) 
%was
%considered to be a favorite for the 110 meter hurdle after running 12.84 seconds
%in early June; only 0.04 seconds away from the world record \citep{wa2022preview}.
%Thus 
His disqualification in the final heat by reacting in 0.099 seconds instead of 
0.1 was met
with shock from both Allen and fans alike.  NBC, who was broadcasting the
championships in the United States, later uploaded a video to their YouTube
channel that showed the disqualification; it can be found here:
\url{https://www.youtube.com/watch?v=D6NXTMo-1yM}. The crowd erupted with jeers 
when they found out that Allen was disqualified, angry at the result.  Had Allen 
reacted just 0.001 seconds slower he would have been allowed to compete and been 
spared waiting another two years until the next World Track and Field Championship.

%\eds{final two sentence seems out of place here; consider moving elsewhere?}
%\of{Moved to the end of paragraph 4 of introduction}

%2)
Allen's disqualification has been discussed extensively in online communities
such as \url{www.LetsRun.com}, a website that is part message board and part
news. The message board,
which functions similarly to Reddit, is very active during important running
events, including during the finals of major events such as
the Olympics and the World Track and Field Championships.  However, it is the
news section of the website that generated articles such as \citet{johnson2022data}
and \citet{johnson2022was} from  Robert Johnson, who created and runs 
the LetsRun website.  He argued that there was an error
with the timing equipment at the 2022 Track and Field World Championship and
cited graphs, computed simple statistics such as median reaction times, and 
compared the reaction times at the 2022 World Track and Field Championship to 
other competitions
such as the USA Track and Field Championships and other World Championships.
The comparisons lacked formal statistical analysis, however. 
The alleged 99.95\% chance Allen was wrong \eds{can we use different word 
instead of wrong?  at fault?  However, should that be 0.05\% chance at fault?}
came from the fact that the lowest average reaction times for the men's 100 meter,
men's 110 meter hurdle, and women's 100 meter competitions held between 2003 and 
2022 each \eds{?} occurred at the 2022 World 
Championships.  Johnson claimed that the probability of this
happening was $\frac{1}{13} \cdot \frac{1}{13} \cdot \frac{1}{13} = 0.05\%$
\citep{johnson2022was}.
%\eds{compared the reaction times at the 2022 World Track and Field  Championship 
%to what? Also restate here that no formal statistical analyses have been done.}
%\of{Cleaned this up a little but still a little messy.}

@@

\jy{Could be condensed}
\citet{pilianidis2012start} found
significant differences between reaction times at several world championships
between 1997 and 2009 for the men's 110 meter hurdles.  Another important
conclusion was that reaction times did not decrease over the twelve years
covered by the research in the study \citep{pilianidis2012start}. This 
contradicts the results of \citet{zhang2021correlation}, which indicated that 
year of competition has a significant effect on reaction times.  Their
analysis looked at data from the 2011-2019 World Championships across multiple
sprint events and concluded that from 2011 to 2019, reaction times got
faster. However, not all studies have been in favor of a decrease in the
reaction time barrier. \citet*{brosnan2017effects} argue for increasing 
the reaction time barrier to 0.115 seconds, and cited evidence from European 
Championships and World Championships in their article.  A study looking 
at the 2008 Olympics likewise supported the conclusion that athletes could not
react in as little as 0.1 seconds \citep{lipps2011implications}.  Another study
concluded that visual reaction times to the starting gun may be faster by 0.007
seconds than the IAAF method, which is from force on block starts 
\citep{holmes2018method}.  While this may seem trivial, Allen was disqualified
only by 0.001 seconds. It is also worth noting that one study looking at data 
from various IAAF meets found that the fastest reaction times for men were found
between the ages of 26 and 29 \citep{tonnessen2013reaction}.  This is notable 
because Allen was 27 years old at the 2022 Track and Field World Championship 
and at his physical peak.

@@

%4)
\jy{language needs to be formalized and condensed.}
The objective to this paper is to analyze the reaction times at the 2022 World 
Track and Field Championship to check for any abnormalities, and we approach
this two different ways.  First we model the data, which is
the reaction times of every non-disqualification since the 1999 World 
Championships, using a gamma linear mixed-effects model with a random effect 
accounting for the venue (or year). \eds{are you using covariates?  Data section
implies that you are, but Results section does not} 
Based on the fitted model, we can then estimate the probability of a reaction 
time being below a certain threshold (e.g., 0.1). 
Second, we compare reaction times for athletes who have competed
at additional competitions to the 2022 World Track and Field Championship.  By
comparing reaction times from the same athlete across multiple competitions we 
are able to determine if the 2022 data differed significantly from other 
competitions.
Our investigation into reaction times is important not only because if its
implications for future races, but also because reaction times have
been shown to affect the overall time of the race \citep{delalija2008reaction}.
Thus, faster reaction times lead to faster overall times; which is the goal of
every sprinter who steps onto the track.
%\eds{I recommend moving the rest of this paragraph to Section 2.2 as I don't
%think it is critical for readers to know all the details here.}
%First we can compare the reaction times of athletes who
%competed at the World Track and Field Championship in both 2019 and 2022, and 
%second we can compare how athletes performed at the 2022 championship relative 
%to their national/qualifiers meet. In order to reduce any chances
%of athletes improving reaction time, we can also look at data from qualifying meets.
%Almost every athlete had to qualify by proving they are elite at the national level, 
%usually by coming in at least fourth. Thus, data was compiled on athletes who competed
%at the 2022 World Track and Field Championships on what these athletes also ran 
%at their country's national championship. If there was any argument that athletes
%have improved their reaction time from 2019 to 2022, it should now be considered
%nearly null, as the national competitions took place between June and July 2022 
%and the World Championships were in August. The equipment used to measure reaction
%time cannot be guaranteed to be the same from national to international
%competition, but the time span between meets has been reduced considerably.

The rest of this paper is organized as follows. Section~\ref{sec:Data} describes 
how data was collected and begins to detail the generalized linear mixed-effect 
(GLMM) model that is further developed in Section~\ref{sec:Methods}. The 
signed-rank test that analyzes reaction times for the same athletes across
different competitions is
also described in Section~\ref{sec:Methods}.  Section~\ref{sec:Results}
 provides the results of the two analyses and ends 
by questioning the reaction time barrier.  Lastly, Section~\ref{sec:Discussion}
discusses the impact and limitations of the paper.
\eds{capitalize first letter when referring to a numbered section, table, figure, 
etc. ``Section 4 provides...'' vs. ``The next section provides...''}

@@

There are two types of data that are explored throughout this paper: data from
1999-2022 for every world championship, and data from athletes who competed in
the 2022 World Championships and another competition (either 2019 World 
Championship or a 2022 national competition held within two months of the  
2022 World Championships). The first data set is used for the GLMM portion
of the paper, while the second data set is used for the rank-based comparisons.

@@

Data was taken from the World Athletics 
\eds{is there a website you can cite here?}
and covers the men's 110 meter hurdles 
and the women's 100 meter hurdles from 1999 to 2022.  The dependent
variable interest is reaction time, and the independent variables are year, 
stage of competition, total time, and gender.  \eds{see earlier comment - 
are you using any of these as covariates?}
The data can be found in the appendix~\ref{sec:Appendix}. Factors such as name, 
country, and identification number were not considered important for the 
purposes of this research besides the times for Devon Allen and the other 
United States athletes mentioned above. \eds{not sure which other US athletes 
you are referring to in the previous sentence} 
The variable ``Stage" refers to whether the observation occurred during a 
%heat (preliminary round)  
preliminary heat denoted by ``H", semifinal heat (of which there are either two
or three each year) denoted by ``S", or final heat denoted by ``F". The variable 
``Gender" has levels ``M" and ``F" for ``male" and ``female" respectively.  The
variable ``Batch" takes a 
numerical value and every heat of every round from 1999 to 2022 was assigned it's
own number.  This was added because of suspected issues that reaction time may be
correlated with other runners reacting quickly (one runner reacting quickly may
result in other runners reacting quickly). There is no statistical meaning to
how the batches were labeled, they were ordered chronologically by meet but
reverse chronologically by year (Batch 1 is a preliminary heat of 2022 and Batch
111 corresponds to the 1999 finals). %Included are
%box plots of the reaction times of the men's finals and the men's semifinals and
%finals pooled together.  
For analysis purposes, we will often pool reaction times of men's final and 
men's semifinal heats together and refer to this data as the ``pooled data"
%We will refer to the latter as the ``pooled data"
throughout the paper.  %The heats or preliminary rounds were not studied because
The preliminary heats were not studied because athletes react faster as the 
competition progresses \citep[e.g.,][]{zhang2021correlation}. % Athletes react 
%faster in the finals than the semifinals which are faster than the heats 
%\citep{zhang2021correlation}.  
\eds{if you are really not using ``H'' data at all, it does not appear necessary
to spend time defining earlier.  You could just say directly when you define 
stage that you only consider semifinals and finals and cite the Zhang paper for 
rationale}
If we want to determine if Allen's reaction time
was reasonable, then we thus need to look at the men's final data, with the men's
pooled data being included in order to increase the sample size.  
\eds{previous sentence is not clear; pooled data already includes final data}
In 2022, there
were only five data points due to two disqualifications and one athlete who 
qualified but did not compete.
The data under consideration is best visualized using side-by-side box plots, as
shown in Figure~\ref{fig:Boxplots}.  It is clear that the 2022 box plot for both 
the men's finals and semifinals data is lower (indicating faster reaction times)
than in previous years.  The median for the men's pooled data in 2022 was 0.129
and the median for the men's final data in 2022 was 0.126.  For comparison,
\citet{brosnan2017effects} found reaction times of 0.156 seconds \eds{mean?  median?} for men when looking at data from 1999 to 2014.

\eds{I wonder if this paragraph would be better placed in the introduction?}
Something that is extremely important to note in the data is the inconsistency 
of World Athletics in terms of how they penalize false starts.  From 2007--2009,
the IAAF (the former name for World Athletics, and the governing body for the 
World Track and Field Champions) instituted a rule change that allowed one 
warning false start before a sprinter was disqualified 
\citep{iaaf2009falsestart}.  This 
essentially gave sprinters a second chance and reduced the penalty for a false 
start.  There were 18 male and 7 female false starts at the 2007 World 
Championships, and 18 male and 7 female starts at the 2009 World Championships. 
Starting in 2011, the rule was scrapped, and the old policy which automatically 
disqualifies runners who false start was reinstated. By returning to the harsher 
policy and cracking down on false starts, World Athletics had reduced the number 
of false starts by two thirds \eds{for men, not overall} in 2011 (6 male and 4 female disqualifications 
\citep{iaaf2009falsestart}).
%The effect? At the 2011 
%World Championship there were 6 male disqualifications and 4 female 
%disqualifications \citep{iaaf2009falsestart}.  
%One study that looked at different false start rules that IAAF has
%imposed from 1997 to 2009 and found that the reaction times when there was less 
%leniency improved by a statistically significant amount
%\citep{haugen2013effect}. 
\citet{haugen2013effect} examined the effect of different false start
rules that IAAF has imposed from 1997 to 2009 and found statistically 
significant improvement in reaction times under harsher rules.
\eds{is that right?  it seems counterintuitive to me}
This was desirable for World Athletics as false starts can make already long 
track meets more tedious, both for athletes and viewers watching the 
television broadcast.  
\eds{I am not sure the previous sentence follows from the one preceding it. Why
 does World Athletics find faster reaction times desireable?  Do you mean that
they find fewer false starts desirable?}
Since 2011, IAAF has instituted the one false start rule
\citep{iaaf2009falsestart}. However, we chose not to include women's data in this
paper due to the numerous papers citing difference in reaction times in men and
women's reaction times: \citep{lipps2011implications, babicc2009reaction,
panoutsakopoulos2020gender}  \eds{Two comments about the previous sentence: 1. 
it does not logically follow the previous sentence as there is no mention of 
gender, and 2. if you are not using women's data, don't spend time talking about 
it at the beginning of the data section.  Similar to my comment about 
preliminary heats, you can move this sentence to the beginning of the Data 
section to provide a rationale for why you do not include data for women.}

\subsection{Beyond World Championship}
The motivation for looking beyond World Championship data came from an 
initial qualitative analysis examining how United States athletes reacted at the 
2022 United States Track and Field Championships.
\eds{be consistent with how you refer to the US championship:
``United States Track and Field Championships" and ``USA Track and Field 
Championships" are both being used; please pick one and use that throughout}
%This section was motivated from a qualitative analysis that started by looking
%at how United States athletes reacted at the 2022 United States Track and Field
%Championships. 
From June 23-26 2022, the United States held its Track
and Field Championship at Hayward 
Field in Eugene, Oregon to decide who to send to the World Championships 
being held at the same venue in August. %the exact same venue as used 
%for the World Track and Field Championship a month later.  
%Thus, a baseline is able to be established for the four USA athletes who 
Thus, we can establish a baseline for the four USA athletes who 
competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
Grant Holloway, and Devon Allen. Every athlete reacted faster in all the World 
Track and Field Championship races compared to the USA Track and Field 
Championship races. Devon Allen raced three times at the USA Track and
Field Championship and three times at the World Track and Field Championship. 
His reaction times were significantly faster in the World Track and
Field Championship compared to the USA Track and Field Championship as were his
teammates. \eds{statistically significantly faster? If you used a statistical 
test, it should not be described in this section.  If you didn't use a 
statistical test, you should replace the word `significantly' with something 
else (substantially?)} 
This comparison suggests that the timing system used between the two
meets may have been meaningfully different to the point where Devon Allen was
disqualified more because of faulty equipment rather than because he reacted
too quickly.  Table~\ref{fig:USAvsWorld} highlights the difference in reaction 
times for Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. 
%``USA" denotes
%the USA Track and Field Championships, ``World" denotes the World Track and Field
%Championships. All numbers listed in the table are the reaction
%time for the respective athletes in seconds.

\begin{table}
\begin{center}
  \caption{Reaction times (in seconds) for four United States Athletes at the 
	2022 USA Track
  and Field World Championships (``USA") and the 2022 World Track and Field 
	Championships (``World"); H denotes preliminary heat, S denotes semifinal, and 
	F denotes final.
  ``N/A" was used for any athlete that did not compete and thus did not register 
  a reaction time. }
  \begin{tabular}{c c c c c c c} 
   \toprule
	 & \multicolumn{3}{c}{USA} & \multicolumn{3}{c}{World} \\
	\cmidrule(lr){2-4}
    \cmidrule(lr){5-7}
   Athlete &  H &  S &  F &  H &  S &  F \\ [0.5ex] 
   \midrule
   Devon Allen & 0.201 & 0.153 & 0.160 & 0.123 & 0.101 & 0.099 \\ 
   Trey Cunningham & 0.186 & 0.185 & 0.182 & 0.115 & 0.120 & 0.109 \\
   Grant Holloway & 0.192 & 0.190 & N/A & 0.147 & 0.128 & 0.124 \\
   Daniel Roberts & 0.181 & 0.200 & 0.183 & 0.179 & N/A & N/A \\ [0.5ex]
   \bottomrule
  \end{tabular}
  \label{fig:USAvsWorld}

@@

This dataset is too small to perform any meaningful statistical analysis, but
the data was expanded to include reaction times for athletes who competed at 
national competitions (many of which were held from May-July of 2022). Once the 
data was collected, each athlete was designated as a cluster and the stage of
competition (national or international) was specified for each competition 
within the cluster (athlete).

@@

The World Championship data and the data beyond the World Championship were
analysed with GLMM and rank-based approaches, respectively.
\eds{need consistency with ``Championship" or ``Championships - I am not sure 
which one is correct}
